Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-10-02 Observations_SAM_SS3_Krill_481.Rmd
changed lot

# Crear el plot usando ggplot
ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "True Age",
       y = "Length",
       fill = "Value") +
  theme_bw()+
  theme(legend.position = "none")
```

@@

kable(data, format = "markdown", 
      col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
      caption = "Table 1: Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
  kable_styling(full_width = FALSE)
```

@@

\newpage
```{r eval=TRUE,, parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

@@

>>>>

# Crear el plot usando ggplot
ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "Age",
       y = "Length",
       fill = "Value") +
  theme_bw()+
  theme(legend.position = "none")
```

@@

kable(data, format = "markdown", 
      col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
      caption = "Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
  kable_styling(full_width = FALSE)
```

@@

\newpage
```{r parainit, eval=TRUE}

@@

And total Harvest rate could be noted in Table \@ref(tab:hrt) and Harvest Rate by Strata regarding outputs of `SS3` in Table \@ref(tab:hrt2)

```{r hrt}
# Calcular el acumulado anual de F_std por año
data <- exploitation_long %>%
  dplyr::filter(Yr<2021) %>% 
  group_by(Yr) %>%
  dplyr::summarize(F_std_accumulated = mean(F_std)) %>% 
  dplyr::select(Yr,  F_std_accumulated) %>% 
  dplyr::rename("Harvest Rate" ="F_std_accumulated" ,
                "Year" = "Yr")

data %>%
  kbl(booktabs = TRUE,
      format = "latex",
      position="ht!",
  caption = "Accumulate Harvest Rate by year in whole 48.1 Subarea") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped"),
                full_width = FALSE,
                font_size=9)

```{r hrt2}
data2 <- base.model1$exploitation %>% 
  dplyr::select(1, 7:11) %>% 
  dplyr::rename("Year" = "Yr")
data2 %>% 
  kbl(booktabs = TRUE,
      format = "latex",
      position="ht!",
      digits = 3,
  caption = "Harvest Rate by strata and year") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped"),
                full_width = FALSE,
                font_size=9) %>% 
  pack_rows("Forecast", 44, 44)
```

### 2024-10-02 SS3_Krill_481_FSA.Rmd
changed lot

The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill (*Euphausia superba*) in an integrated stock assessment model. Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making. The analysis represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. We have three main objectives. i)integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance; ii) to consider the spatial heterogeneity of the krill population structure and, iii) to evaluate the impact of biological and  population structure assumptions on the performance of this type of integrated stock assessment models. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management strategies in this region. This study addresses key recommendations provided by the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) . The base model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions. Retrospective analysis showed a slight underestimation bias in spawning biomass; however, this was within an acceptable range. The model suggests a higher dependence on length composition data compared to other sources for estimating recruitment. Spatial considerations and assumptions about stock-recruitment relationships significantly impact population variable estimates. The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the base model (`s1`). Overall, this study demonstrates the potential of using a spatially explicit integrated model for krill stock assessment. The model provides valuable insights into krill population dynamics and can be used to inform sustainable management practices.

@@

Here we have incorporated a scenario (`s9`) with a selectivity block for krill. The inclusion of selectivity blocks aims to reduce the effect of autocorrelation in the model fitting process, and also, to explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms. According to the data identified, a shift from traditional to continuous trawling occurred in 2006. As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements. This is compared with the base scenario.

@@

```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
knitr::include_graphics('Figs/recipediags.png')
```

@@

To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding these configurations. We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).

@@

## Diagnosis Base Model

@@

A selectivity block in an alternative scenario (`s9`) related trawl thecnics and mesh changes was implemented and compared its performance with the base model (`s1`). The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)). No significant differences were detected in the state variables.

>>>>

The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill (*Euphausia superba*) in an integrated stock assessment model. Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making. The analysis represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. We have three main objectives. i)integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance; ii) to consider the spatial heterogeneity of the krill population structure and, iii) to evaluate the impact of biological and  population structure assumptions on the performance of this type of integrated stock assessment models. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management strategies in this region. This study addresses key recommendations provided by the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) . The reference model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions. Retrospective analysis showed a slight underestimation bias in spawning biomass; however, this was within an acceptable range. The model suggests a higher dependence on length composition data compared to other sources for estimating recruitment. Spatial considerations and assumptions about stock-recruitment relationships significantly impact population variable estimates. The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the reference model (`s1`). Overall, this study demonstrates the potential of using a spatially explicit integrated model for krill stock assessment. The model provides valuable insights into krill population dynamics and can be used to inform sustainable management practices.

@@

Here we have incorporated a scenario (`s9`) with a selectivity block for krill. The inclusion of selectivity blocks aims to reduce the effect of autocorrelation in the model fitting process, and also, to explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms. According to the data identified, a shift from traditional to continuous trawling occurred in 2006. As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements. This is compared with the reference scenario.

@@

```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a reference model."}
knitr::include_graphics('Figs/recipediags.png')
```

@@

To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding these configurations. We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the reference model (Table 1).

@@

## Diagnosis Model

@@

A selectivity block in an alternative scenario (`s9`) related trawl thecnics and mesh changes was implemented and compared its performance with the reference model (`s1`). The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)). No significant differences were detected in the state variables.

### 2024-10-02 index.Rmd
changed lot

```{r}
parInit %>%
  kbl(booktabs = T,
      format = "html",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9)%>% 
  pack_rows(index = c("Mortalidad natural" = 1,
                        "Crecimiento"= 5,
                        "Relación longitud-peso" = 2,
                        "Ojiva de madurez"=2,
                        "Relación stock-recluta"=5,
                        "Capturabilidad"=2,
                        "Selectividad"=4))

@@

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir9,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

```{r eval= FALSE, message=F, include=FALSE}
# html
SS_plots(base.model9,
         uncertainty=T,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75,
         minbthresh=0.20, 
         forecast=T)

@@

```{r fig.height=6}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.model7,
            subplots = 1)
```
```{r}
SSplotComps(base.model8,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

@@

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps1 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
  kable_paper("hover", 
              full_width = TRUE)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9)#%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

@@

>>>>

```{r}
parInit %>%
  kbl(booktabs = T,
      format = "html",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=10,
                html_font = "arial")%>% 
  pack_rows(index = c("Mortalidad natural" = 1,
                        "Crecimiento"= 5,
                        "Relación longitud-peso" = 2,
                        "Ojiva de madurez"=2,
                        "Relación stock-recluta"=5,
                        "Capturabilidad"=2,
                        "Selectividad"=4))

@@

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

```{r eval= FALSE, message=F, include=FALSE}
# html
SS_plots(base.model1,
         uncertainty=T,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75,
         minbthresh=0.20, 
         forecast=T)

@@

```{r fig.height=6}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.model7,
            subplots = 1)
```
```{r}
SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

@@

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps1 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
  kable_paper("hover", 
              full_width = TRUE)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9,
                html_font = "arial")#%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

@@

In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process
